### 2021-09-17 epitweetr_setup.Rmd

---
title: "epitweetr"
author: "C. Coleman"
date: "9/13/2021"
output: html_document
---

#*Set-up*
#Checking if Java Home is in the System environment. Make sure it doesn't return "NULL".
```{r}
Sys.getenv("JAVA_HOME")
```

#You have to go online and download Rtools to run epitweetr. Once downloaded, install these packages.
```{r}
install.packages(DT, type = "source")
```

#Load epitweetr.
```{r}
library(epitweetr)
```

#opening epitweetr dashboard and making quere for covid-19 searches.
```{r}
epitweetr::epitweetr_app("data_dir")
```
#*Twitter Authentification*
#Install rtweet package. 
```{r}
install.packages(rtweet, type = "source")
```
#Load rtweet
```{r}
load(rtweet)
```