### 2021-11-12 results/Supplementary_Data.Rmd
Supplementary_Figures - CC

---
title: "Supplementary Data"
author: "C. Coleman"
date: "11/12/2021"
output: html_document
---

_The following markdown serves as the supplementary materials for the UOWN Stream Health Indicators Project. Included materials are all figures relevant to my manuscript and were derived from the analysis code._

</br>

_Load needed libraries_
```{r}
library(knitr)    # For knitting document and include_graphics function
library(ggplot2)  # For plotting
library(png)      # For grabbing the dimensions of png files
```

</br>

_Figure 1: Plots of biological score versus conductivity, E. coli cfu, nitrate concentration, and turbidity for the MIDO site._
```{r}
bs_MIDO_location <- here::here("results", "MIDO_bs_lm.png")

include_graphics(bs_MIDO_location)
```

</br>

_Figure 2: Plots of E. coli cfu versus conductivity, nitrate concentration, and turbidity for the MIDO site._
```{r}
ecfu_MIDO_location <- here::here("results", "MIDO_ecfu_lm.png")

include_graphics(ecfu_MIDO_location)
```

</br>

_Figure 3: Plots of biological score versus conductivity, E. coli cfu, nitrate concentration, and turbidity for the NORO site._
```{r}
bs_NORO_location <- here::here("results", "NORO_bs_lm.png")

include_graphics(bs_NORO_location)
```

</br>

_Figure 4: Plots of E. coli cfu versus conductivity, nitrate concentration, and turbidity for the NORO site._
```{r}
ecfu_NORO_location <- here::here("results", "NORO_ecfu_lm.png")

include_graphics(ecfu_NORO_location)
```

</br>

_Figure 5: Plots of biological score versus conductivity, E. coli cfu, nitrate concentration, and turbidity for the BICO site._
```{r}
bs_BICO_location <- here::here("results", "BICO_bs_lm.png")

include_graphics(bs_BICO_location)
```

</br>

_Figure 6: Plots of E. coli cfu versus conductivity, nitrate concentration, and turbidity for the BICO site._
```{r}
ecfu_BICO_location <- here::here("results", "BICO_ecfu_lm.png")

include_graphics(ecfu_BICO_location)
```